Model

### 2019-04-21 stats/model.Rmd
inverse model predictions

```{r}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

@@

@@

  var so_slope = uniformDrift({a: -2, b: 2, width: 0.2})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.2})

@@

  query.add("speaker_optimality", [so_int, so_slope])
  return query 
}
'
```

@@

# sem_know_model<- webppl(
#   program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
#   data = list(data = me_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 900, burn = 300, verbose = T)
# )
#  
# saveRDS(sem_know_model, "../saves/sem_know_model.rds")

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH", 
                        samples = 5000, 
                        burn = 2500,
                        verbose = T,
                        verboseLag = 500)
)
```

@@

sem_know_params <- sem_know %>%
  group_by(item, parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))
```

```{r}
ggplot(sem_know, aes(value))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(parameter~item, scales = 'free')+
  theme_few()
```
## Plotting semantic knowledge
```{r}

aoa_ratings_all <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)

@@

@@

age_me_effect <- as_tibble(ranef$familiar, rownames = "familiar")%>%
  mutate(grand_intercept = fixef%>%filter(term=="Intercept")%>%pull(Estimate),
         grand_slope = fixef%>%filter(term=="z_age")%>%pull(Estimate))%>%
  group_by(familiar) %>%
  expand(Estimate.Intercept,Estimate.z_age,grand_intercept,grand_slope,z_age = me_data$age_num - mean(me_data$age_num), mean_age = mean(me_data$age_num))%>%
  mutate(me_effect = plogis(grand_intercept + Estimate.Intercept+(Estimate.z_age+grand_slope)*z_age), 
         age = z_age+mean_age)%>%
  select(familiar, me_effect, age)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(familiar = fct_reorder(factor(familiar), mean_aoa))

model_plot <- ggplot(age_me_effect, aes(x=age,y = me_effect, col = familiar))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Model")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")

p2 <- read_csv("../data/child_basic_me.csv")%>%
  filter(trial != "train1",
         trial != "train2")%>%
    left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(familiar = fct_reorder(factor(familiar), mean_aoa))

data_plot <- ggplot(data = p2, aes(x = age_num, y = correct, col = familiar, fill = familiar)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  geom_smooth(method = "loess", se = F, alpha = .5)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(~familiar)+
  ggtitle("Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")+
  scale_fill_viridis_d(name = "familiar object")

@@

```{r}
ggarrange(data_plot,model_plot, common.legend = T, legend = "right")
```

>>>>

```{r}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

@@

me_data <- read_csv("../data/child_basic_me.csv")

@@

  var so_slope = uniformDrift({a: -2, b: 2, width: 0.2})
  var so_int = uniformDrift({a: -2, b: 2, width: 0.2})
  //var speakerOptimality = uniformDrift({a: -5, b: 5, width: 0.2})

@@

  //query.add("speaker_optimality", [speakerOptimality])
  query.add("speaker_optimality", [so_int, so_slope])
  return query 
}
'
```

@@

# sem_know_model<- webppl(
#   program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
#   data = list(data = me_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 4,
#   cores = 4,
#   inference_opts = list(method = "MCMC", samples = 2000, burn = 1000, verbose = T)
# )
# 
# saveRDS(sem_know_model, "../saves/sem_know_model.rds")

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH", 
                        samples = 5000, 
                        burn = 2500,
                        verbose = T,
                        verboseLag = 500)
)

#saveRDS(sem_know_model, "../saves/sem_know_model.rds")
sem_know_model <- readRDS("../saves/sem_know_model.rds")
```

@@

sem_know_params <- sem_know %>%
  group_by(item, parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

speak_opt <- sem_know_model %>%
  filter(Parameter == "speaker_optimality") %>%
  separate(value, into = c("intercept", "slope"), 
           sep = ", ")%>%
  mutate(intercept = as.numeric(str_remove_all(intercept, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]"))) %>%
  rename(item = Parameter) %>%
  gather(parameter, value, intercept, slope)

```{r}
ggplot(sem_know, aes(value))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(parameter~item, scales = 'free')+
  theme_few()
```

```{r}
ggplot(speak_opt, aes(value))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(~parameter, scales = 'free')+
  theme_few()
```

## Plotting semantic knowledge
```{r}
aoa_ratings_all <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)

@@

# inverse predictions

```{r}
speak_opt_params <- speak_opt %>%
  group_by(parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

inv_model_data <- sem_know_model %>%
  filter(Parameter != "speaker_optimality") %>%
  mutate(Parameter = factor(Parameter)) %>%
  separate(value, into = c("intercept", "slope"), 
           sep = ", ") %>%
  mutate(intercept = as.numeric(str_remove_all(intercept, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]"))) %>%
  rename(item = Parameter)%>%
  group_by(item)%>%
  summarise(sem_int = estimate_mode(intercept),
            sem_slope = estimate_mode(slope))%>%
  group_by(item)%>%
  expand(sem_int,sem_slope,age = me_model_data$age)%>%
  mutate(so_int = speak_opt_params%>%filter(parameter == "intercept")%>%pull(mode),
         so_slope = speak_opt_params%>%filter(parameter == "slope")%>%pull(mode))%>%
  mutate(speaker_optimality = so_int + so_slope * age,
         semantic_knowledge = plogis(sem_int + sem_slope * age))

```{r}
invPred <- '
var allData = dataFromR.data

var output = map(function(row){

    var modelPredictions = pragmaticListener({label: "novel_word"}, priorProbs, row.speaker_optimality,row.semantic_knowledge)

    return extend([row.item + "/" + row.age, Math.exp(modelPredictions.score(1))])

}, allData)

output

```{r predictions pragmatic model ex 3 adults}
inv_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, invPred , sep='\n'),
  data =list(data = inv_model_data),
  data_var = "dataFromR"
)

#saveRDS(inv_model_pred, "../saves/inv_model_pred.rds")
#inv_model_pred <- readRDS("../saves/inv_model_pred.rds")

inv_predictions <- inv_model_pred %>%
  separate(`0`, into = c("item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa), 
         age = age + min(me_data$age_num))

inv_model_plot <- ggplot(inv_predictions, aes(x=age,y = me_effect, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Inverse Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()
```

@@

age_me_effect <- as_tibble(ranef$familiar, rownames = "familiar")%>%
  mutate(grand_intercept = fixef%>%filter(term=="Intercept")%>%pull(Estimate),
         grand_slope = fixef%>%filter(term=="z_age")%>%pull(Estimate))%>%
  group_by(familiar) %>%
  expand(Estimate.Intercept,Estimate.z_age,grand_intercept,grand_slope,z_age = me_data$age_num - mean(me_data$age_num), mean_age = mean(me_data$age_num))%>%
  mutate(me_effect = plogis(grand_intercept + Estimate.Intercept+(Estimate.z_age+grand_slope)*z_age), 
         age = z_age+mean_age,
         item = familiar)%>%
  select(item, me_effect, age)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

model_plot <- ggplot(age_me_effect, aes(x=age,y = me_effect, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("GLMM")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")

p2 <- read_csv("../data/child_basic_me.csv")%>%
  filter(trial != "train1",
         trial != "train2")%>%
  mutate(item = familiar)%>%
    left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(familiar = fct_reorder(factor(item), mean_aoa))

data_plot <- ggplot(data = p2, aes(x = age_num, y = correct, col = familiar, fill = familiar)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  geom_smooth(method = "loess", se = F, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(~familiar)+
  ggtitle("Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")+
  scale_fill_viridis_d(name = "familiar object")

@@

```{r}
ggarrange(data_plot,inv_model_plot,model_plot, common.legend = T, legend = "right", nrow = 1, ncol = 3)
```
